int-theory-vacc: Study 2 modeling

### 2018-09-20 study2/cog-model-discrete-predictions.Rmd
minor readme update

print(Sys.time() - t1)
```

@@

>>>>

print(Sys.time() - t1)

saveRDS(object = boot_fits, file = "boot_fits.rds")
```

@@

# two beliefs at the same time

```{r}
res_overpar <- find_evid_ratio(filter_dscored(d_scored, "overpar"))
res_ds <- find_evid_ratio(filter_dscored(d_scored, "diseaseSevere"))

overpar_ratio <- exp(res_overpar$par[1] + res_overpar$par[2])
ds_ratio <- exp(res_ds$par[1] + res_ds$par[2])

p0p0 <- .3
p1p0 <- p0p0*ds_ratio
p0p1 <- p0p0*overpar_ratio
p1p1 <- p0p0*overpar_ratio*ds_ratio

overpar_ds_evid_cpt <- make_custom_cpt(
  # c(p1p0, p0p0, (1-p1p0), (1-p0p0), p1p1, p0p1, (1-p1p1), (1-p0p1)),   
  c(p1p1, (1-p1p1), p1p0, (1-p1p0), p0p1, (1-p0p1), p0p0, (1-p0p0) ),
                c(2,2,2),
                list(evid=c("Yes","No"), overpar = c("Yes","No"), diseaseSevere = c("Yes","No")), 
                NULL)

```{r}
add_evid_node_custom <- function(bnlearn_model, parents_string, evid_cpt) {

  # evid_cpt <- make_evid_cpt(data, targetted_belief)
  model_list <- lapply(bnlearn_model, function(x){x$prob})
  # model_list$evid <- hyde_to_bn_cpt(evid_cpt)
  model_list$evid <- evid_cpt

  model_string <- modelstring(bnlearn_model)
  model_string <- paste0(model_string,"[evid|", parents_string, "]")
  model_network <- model2network(model_string)
  predModel <- custom.fit(model_network, dist = model_list)

  return(predModel)

ds_overpar_model <- add_evid_node_custom(base_model, "overpar:diseaseSevere", overpar_ds_evid_cpt)

# ----------------------------------------

set.seed(12345)

pred0_ds_overpar <- sapply(nodes, function(x){
  statement <- paste0("cpquery(ds_overpar_model, event = (", x, " == 'Yes'), evidence=TRUE, n = 5e5)")
  eval(parse(text=statement))
})

pred1_ds_overpar <- sapply(nodes, function(x){
  statement <- paste0("cpquery(ds_overpar_model, event = (", x, " == 'Yes'), evidence= (evid == 'Yes'), n = 5e5)")
  eval(parse(text=statement))
})

pred_changes2 <- pred1_ds_overpar - pred0_ds_overpar

pred_changes2 <- as.data.frame(pred_changes2)

pred_changes2 <- pred_changes2 %>%
  mutate(scale = nodes) %>%
  rename(Mean = pred_changes2) %>%
  mutate(type = "predicted")

# ----------------------------------------

obs_pred2 <- bind_rows(pred_changes2, obs_changes) %>%
  spread(type, Mean) %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE)

cor(obs_pred2$predicted, obs_pred2$observed)

obs_pred2 %>% # a bit of a hack
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_point() +
  geom_errorbar() +
  coord_cartesian(xlim=c(-.06,.08), ylim=c(-.06,.08)) +
  theme_bw() +
  theme(aspect.ratio = 1)

obs_pred2 %>%
  gather(type, Mean, observed, predicted) %>%
  ggplot(aes(y = reorder(scale, Mean), x = Mean, xmin = ll, xmax = ul, color=type)) +
  geom_point() +
  geom_errorbarh(height=0)
```

```{r}
# obs,pred correlation excluding nodes fit
obs_pred_excl <- bind_rows(pred_changes, obs_changes) %>%
  spread(type, Mean) %>%
  filter(scale!="diseaseSevere", scale!="overpar") %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE)

cor(obs_pred_excl$predicted, obs_pred_excl$observed)
sqrt(mean((obs_pred_excl$observed-obs_pred_excl$predicted)^2))
```

```{r}
# obs,pred correlation excluding nodes fit
obs_pred_excl2 <- bind_rows(pred_changes2, obs_changes) %>%
  spread(type, Mean) %>%
  filter(scale!="diseaseSevere",scale!="overpar") %>%
  group_by(scale) %>%
  summarize_all(mean, na.rm=TRUE)

cor(obs_pred_excl2$predicted, obs_pred_excl2$observed)
sqrt(mean((obs_pred_excl2$observed-obs_pred_excl2$predicted)^2))
```

Adding edge from overparenting hurts prediction for the 12 nodes other than disease severity and overparenting, goes from r = .90 to r = .818. RMSE just about doubled. To be sure I'd have to do the bootstrap on these estimates, but I think this is a signal that the fit is worse modeling it as a direct effect. If so that indicates to me that something is wrong in the model wrt overparenting beliefs.
